---
title: "Lab 09: Factors, Dates and Times"
author: Owen G. Ward
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r}
library(tidyverse)
library(forcats)
library(lubridate)
yvr <- read_csv("weatherYVR.csv",
  col_types = cols(
    `Date/Time` = col_character()
  )
)
yvr
```

Run the above code chunk to read in **daily** weather
data from YVR in 2003. Coercing `Date/Time` to a character vector is done because different versions of R and operating systems may read it in as a date-time object instead, negating Exercise 1 below.

1. Coerce the `Date/Time` variable to a date object
and rename it Date.

```{r}

```

2. Make a time series plot (with lines) 
of the daily maximum temperature by day.

```{r}

```

3. Change the `Month` variable from 
numeric to a factor.
(Hint: The `month()` function with the `label=TRUE` argument will
extract the months from a date-time object.)

```{r}


```

4. Plot the average maximum temperature *versus* month.
Then, redo this plot with months ordered by 
average maximum.

```{r}



```

```{r}


```
