---
title: "Lab 08: Relational Data - Solution"
author: Owen G. Ward
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```


```{r}
library(tidyverse)
library(nycflights13)
```

In this lab we will work with the `nycflights13` data.

1. Add the latitude and longitude of each 
airport destination to the `flights` table using a `join` function.
You will find the data on latitude and longitude
in the `airports` table.
```{r}
head(flights)
head(airports)
```

```{r}

flights_with_coords <- flights %>%
  left_join(airports, by = c("dest" = "faa"))

# View the result
head(flights_with_coords)

```

2. Create a table with the
year-month-day-flight-tailnum
combinations that have more than 1 flight (careful about
missing tailnum). Use this table to filter the flights
table and then select carrier, flight, origin and dest.
Which airline used the same flight number for a plane
that made a trip from La Guardia to St. Louis in 
the morning and from Newark to Denver in the afternoon?

```{r}
duplicate_flights <- flights %>%
  group_by(year, month, day, flight, tailnum) %>%
  filter(n() > 1) %>%
  ungroup()

filtered_flights <- flights %>%
  semi_join(duplicate_flights, by = c("year", "month", "day", "flight", "tailnum")) %>%
  select(carrier, flight, origin, dest)

head(filtered_flights)
morning_flight <- filtered_flights %>%
  filter(origin == "LGA" & dest == "STL") 
afternoon_flight <- filtered_flights %>%
  filter(origin == "EWR" & dest == "DEN")

common_flights <- morning_flight %>%
  inner_join(afternoon_flight, by = "flight", suffix = c(".morning", ".afternoon"))

common_flights
```

3. One of the exercises in the lecture 6 notes asked you
to create a table called `top_dep_delay` from
the flights table. 
`top_dep_delay` was
comprised of the year-month-days with the 
3 largest total delays, where total delay
is defined as the sum of the `dep_delay` variable
for each year-month-day. 
Recreate `top_dep_delay` for this lab exercise.
For each of the three top-delay days, 
report the median, third quartile and maximum
of the dep_delay variable in `flights`.

```{r}
top_dep_delay <- flights %>%
  group_by(year, month, day) %>%
  summarise(total_dep_delay = sum(dep_delay, na.rm = TRUE)) %>%
  arrange(desc(total_dep_delay)) %>%
  slice_head(n = 3)

print(top_dep_delay)

summary_stats <- flights %>%
  filter(paste(year, month, day) %in% paste(top_dep_delay$year, top_dep_delay$month, top_dep_delay$day)) %>%
  summarise(
    median_dep_delay = median(dep_delay, na.rm = TRUE),
    third_quartile_dep_delay = quantile(dep_delay, 0.75, na.rm = TRUE),
    max_dep_delay = max(dep_delay, na.rm = TRUE)
  )

print(summary_stats)
```
