---
title: "Lab 04: Tidy Data"
author: Your Name
format: html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## HIV prevalence from WHO

* We used a tidy version of the HIV prevalence data in Lab 2, and
saw the raw version in Lab 3. In this lab we will tidy the latter into the former.

Don't worry about the use of `read_csv` below, we will discuss this next 
week.

```{r,message=TRUE}
library(tidyverse)
hiv <- read_csv("HIVprevRaw.csv")
hiv
```

(The columns for 1988 and 1989 are completely empty and 
were read in as logical. We will be removing these and so won't
worry about over-riding the logical with double.)


For each of these tasks below you should create a new object 
with the modified version of the `hiv` data. Use a consistent
and informative naming convention to make your life easier.

1. The first column of the data frame is the country, but it has been named`Estimated HIV Prevalence% - (Ages 15-49)`.
Use the `rename()` function to rename this column  `Country`.
(Hint: The current variable name contains special characters.)

```{r}
hiv <- rename(hiv, Country = 'Estimated HIV Prevalence% - (Ages 15-49)')
```


2. The data from 1979 to 1989 is very sparse. Remove these columns from the data frame.

```{r}
hiv <- select(hiv, -('1979':'1989'))
```

3. Pivot the yearly prevalence estimates into a longer tibble that contains only three columns: `Country`, `year`, and `prevalence`. When you pivot,
remove explicitly missing values. After pivoting, sort the resulting tibble by `Country`. 

```{r}
hiv |> 
  pivot_longer(
    cols = c('1990' : '2011'),
    names_to = "year",
    values_to = "prevalence",
    values_drop_na = TRUE
  ) |>
  arrange(Country,year)
```
