---
title: "Lab 03: Data Transformation"
author: haoran Miao
format: html
---



```{r setup, include=FALSE, message=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#| message: false

## first load in the tidyverse packages
library(tidyverse)
```



## HIV prevalence from WHO

- The HIV prevalence data from Lab 2 was modified from its raw form. 
In this lab we will work with the raw data to explore data transformation.

- A spreadsheet of the data `HIVprevRaw.csv` is available in 
the Labs-->Lab 3 folder on Canvas. We 
can read in these data as follows (we'll learn about reading in data in later 
weeks; note that you must have `HIVprevRaw.csv` in your current working 
directory or supply the full path in `read_csv()`; 
please consult your TA if you encounter difficulty):

```{r}
hiv <- read_csv("HIVprevRaw.csv")
```

- Take a look at the top few lines of raw data.

```{r}
head(hiv)
```

- Make a copy of `hiv` for use in exercise 4.

```{r}
hivcopy <- hiv
```


- In exercises 1 - 3, save the results of each data manipulation in 
`hiv`. In exercise 4  you will use the copy `hivcopy`.


**Exercises:**


1. The first column of the data frame is the country, but it
has been named:
`Estimated.HIV.Prevalence.....Ages.15.49.`.
Use the `rename()` function to rename this column  `Country`.

```{r}
hiv <- rename(hiv, Country = 'Estimated HIV Prevalence% - (Ages 15-49)')


```


2. The data from 1979 to 1989 is very sparse. Remove these
columns from the data frame.

```{r}
hiv <- select(hiv, -('1979':'1989'))

```


3. Sort the data in descending order of prevalence in 2011. 
Print the first 6 rows of 
your final data set.

```{r}
hiv <- hiv %>% arrange(desc(`2011`))
head(hiv, 6)

```


4. Use the copy `hivcopy` and the pipe operator to chain or ``pipe'' the data manipulations 
of exercises 1-3.

```{r}

hivcopy |>
  rename(Country = 'Estimated HIV Prevalence% - (Ages 15-49)') |>
  select(hiv, -('1979':'1989')) |>
  arrange(desc(`2011`)) |>
  head(6) 

  

```


