---
title: "Lab 06: Midterm Recap"
author: Owen G. Ward
format: pdf
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r, include=TRUE, message=FALSE}


library(tidyverse)
```



In this lab we will briefly recap some of the key `R` functions you have
seen in the past few weeks. 

1. You should first read in the file provided,
`canada_births.csv`,  using the appropriate `read_`
function,
removing the comments in the file. 
Assign this to an `R` object called `canada_births`. 
Print the first 5 rows of this file. 
This file describes the number of births in Canada each month 
for the last several years.

```{r}

canada_births <- read.csv("canada_births.csv", comment.char = "#")


```


2. This data is not tidy, so we will first tidy it. Use `pivot_longer`
to create a new tibble, `canada_births_tidy` where 
we have the births for a single year and month in each row.
This should create a tibble with columns that
you should call `year,month,births`.

```{r}
canada_births_tidy <- canada_births %>%
  pivot_longer(
    cols = -year,       
    names_to = "month", 
    values_to = "births"
  )
```


3. Plot this data, where `month` is on the horizontal axis,
the number of births is on the vertical axis and there
is a different line for each year in the dataset.
Color these lines red and use `alpha=0.5` to show all
lines. To do this plot you will need to first convert the `month` column 
to be a number, using `mutate` and `parse_number`


```{r}
ggplot(canada_births_tidy, aes(x = month, y = births, group = year)) +
  geom_line(color = "red", alpha = 0.5) + 
  labs(x = "Month", y = "Number of Births") 
 
```



