---
title: "Lab 07: Midterm Recap"
author: Owen G. Ward
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(palmerpenguins)
```

In this lab we will answer some of the questions from the 
midterm using code, to see how they would be solved in the real world.

To complete this lab you will need the `penguins` data which is
in the R package `palmerpenguins`.


1. We will first create a scatterplot of the penguin data, where the
`x` variable is `bill_length_mm`, the `y` variable is 
`bill_depth_mm` and the points are colored by `sex`.

Note you will have some `NA` points here which we don't need to worry about.


```{r}
# Create scatterplot
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point()
```


2. Add a facet to the previous plot, corresponding to the variable `island`.

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point()+
facet_wrap(~ island)
```


3. Finally we will create a more complex plot. 
We want to have a separate smoother for male and femal penguins, 
along with coloring the points by `sex`. Finally, we want to have separate
facet plots for each of the three species. Note you will also get a smoother
for `NA` values which we can ignore. Set `se = FALSE` in the smoother
function to make the plot a bit tidier.

```{r}

ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm, color = sex)) +
  geom_point() +
  geom_smooth(aes(group = sex), method = "loess", se = FALSE) +
  facet_wrap(~ species)

```


